GPU-Beschleunigung der Material Point Method

### 2018-12-08 mscthesis.tex
evaluation start

\documentclass[m,times]{cgMA}
\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english,ngerman]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{mathtools}
\usepackage{subdepth}
\usepackage{enumitem}
\usepackage[toc,page]{appendix}
\usepackage{minted}
\usepackage{verbatim}
\usepackage{blindtext}
\addbibresource{mscthesis.bib}

@@

    \begin{equation}
      \text{Speedup of Procedure 1} = \frac{\Delta t_{c1}}{\Delta t_{p1}} \frac{\Delta t_{p2}}{\Delta t_{c2}}.
    \end{equation}
\end{enumerate}
\begin{table}
  \begin{tabular}{ | l | l | l |}    \hline
    \textbf{Metric}   & \textbf{Max}(GTX 970) & \textbf{Description} \\ \hline
    Speedup           & -                     & item \ref{it:speedup}\\\hline
    $\Delta t_c$      & -                     & computing time\\\hline
    VRAM SOL\%        & 100                   & memory througput(, w.r.t. to hw.-limit)\\\hline
    SM SOL\%          & 100                   & instruction throughput\\\hline
    L2 SOL\%          & 100                   & L2-cache throughput\\\hline
    Tex SOL\%         & 100                   & L1-cache throughput\\\hline
    L2-Hitrate\%      & 100                   & L2 hit-miss ratio\\\hline
    SM Active\%       & 100                   & at least one warp active avg. over SMs\\\hline
    SM Occupancy      & 64.0                  & active warps avg. over SMs\\\hline
    Read Active\%     & 100                   & \% of time spend on VRAM reads\\\hline
    Write Active\%    & 100                   & \% of time spend on VRAM writes\\\hline
    \multicolumn{3}{|l|}{\textbf{Warp Stall Reasons:}}\\\hline
    Long Scoreboard\% & 100                   & memory latency\\\hline
    Barrier\%         & 100                   & block synchronization \\
    %No Instructions\% & 100                   & \\
    \hline
  \end{tabular}
\caption{GPU Performance Metrics}\label{table:metric}
\end{table}
The other GPU metrics are acquired using NVIDIA Nsight with OpenGL Performance Markers \cite{NVIDIA:METRICS}. An active warp is also one that waits for memory.

@@

In short one might not access every variable of a particle in a shader. I.e. in physics one has a pipeline of different passes each requiring another subset of variables. Assume for instance one shader only requires the position then it is more favorable for consecutive threads to load consecutive position vectors. The cache-line is then filled with position vectors, which all will get used. In contrast in an AoS layout the cache-line will have a position and a velocity loaded, but the shader never uses the velocity, halfing the bandwidth.
\noindent\begin{wrapfigure}{R}{0.5\textwidth}
\hfill\begin{minipage}{0.45\textwidth}
\begin{lstlisting}[caption={OpenGL Layout},label={llst:opengl},language={GLSL},style={GL}]
//if SoA
#define Particle PREC_VEC_TYPE
#define Particle_position 0
#define Particle_velocity 1
#define Particle_size 2
//if AoS
struct Particle{
  PREC_VEC_TYPE position;
  PREC_VEC_TYPE velocity;
};
buffer Container{
  Particle variables [ ];
};
\end{lstlisting}
\end{minipage}
\end{wrapfigure}

@@

\subsection{Parallel Reduction}
\cite{NVIDIA:PARALLEL_REDUCTION}
\subsection{Counting sort}
Map (Reset) \\
\subsubsection{Bin}
Binning\cite{NVIDIA:BINNING} (coalesced, uncoalesced)\\
\subsubsection{Scan}
Scan\cite{NVIDIA:SCAN}\\
Scan(Raking) \cite{NVIDIA:SCAN_MODERN}
\subsubsection{Reordering}
Coalesced, Uncoalesced
\subsection{MPM-Transfers}
Conjugate gradient method
\subsubsection{Particle to grid transfers}
\subsubsection{Grid to particle transfers}
\subsection{Verifying Simulation}
\begin{itemize}
  \item Conservation equations
    \cite{MPM:APIC}
  \item Singular Value Decomposition
  \item Conjugate gradient
\end{itemize}
\section{Evaluation}
\cite{NVIDIA:PEAK_PERFORMANCE}
\cite{KHRONOS:TIMER_QUERY}
\section{Conclusion}
\subsection{Review}
\subsection{Future Work}
Particle Activation: \cite{MPM:GPU}

>>>>

\documentclass[m,times]{cgMA}
\usepackage{listings}
\usepackage{xurl}
\urlstyle{sf}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{color}
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage[english,ngerman]{babel}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{mathtools}
\usepackage{subdepth}
\usepackage{enumitem}
\usepackage[toc,page]{appendix}
\usepackage{minted}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{blindtext}
\usepackage{svg}
\addbibresource{mscthesis.bib}

@@

    \begin{equation}
      \text{Speedup of Procedure 1} = \frac{\Delta t_{c1}}{\Delta t_{p1}} \frac{\Delta t_{p2}}{\Delta t_{c2}}.
    \end{equation}
\end{enumerate}
\begin{table}
  \begin{tabular}{ | l | l | l |}    \hline
    \textbf{Metric}   & \textbf{Max}(GTX 970) & \textbf{Description} \\ \hline
    Speedup           & -                     & item \ref{it:speedup}\\\hline
    $\Delta t_c$      & -                     & computing time\\\hline
    VRAM SOL\%        & 100                   & memory througput(, w.r.t. to hw.-limit)\\\hline
    SM SOL\%          & 100                   & instruction throughput\\\hline
    L2 SOL\%          & 100                   & L2-cache throughput\\\hline
    Tex SOL\%         & 100                   & L1-cache throughput\\\hline
    L2-Hitrate\%      & 100                   & L2 hit-miss ratio\\\hline
    SM Active\%       & 100                   & at least one warp active avg. over SMs\\\hline
    SM Occupancy      & 64.0                  & active warps avg. over SMs\\\hline
    Read Active\%     & 100                   & \% of time spend on VRAM reads\\\hline
    Write Active\%    & 100                   & \% of time spend on VRAM writes\\\hline
    \multicolumn{3}{|l|}{\textbf{Warp Stall Reasons:}}\\\hline
    Long Scoreboard\% & 100                   & memory latency\\\hline
    Barrier\%         & 100                   & block synchronization \\\hline
    %No Instructions\% & 100                   & \\
    Drain\%           & 100                   & memory wait before warp exit \\
    \hline
  \end{tabular}
\caption{GPU Performance Metrics}\label{table:metric}
\end{table}
The other GPU metrics are acquired using NVIDIA Nsight with OpenGL Performance Markers \cite{NVIDIA:METRICS}. An active warp is also one that waits for memory.

@@

In short one might not access every variable of a particle in a shader. I.e. in physics one has a pipeline of different passes each requiring another subset of variables. Assume for instance one shader only requires the position then it is more favorable for consecutive threads to load consecutive position vectors. The cache-line is then filled with position vectors, which all will get used. In contrast in an AoS layout the cache-line will have a position and a velocity loaded, but the shader never uses the velocity, halfing the throughput. Chapter \ref{sec:eval}.

\noindent\begin{wrapfigure}{R}{0.5\textwidth}
\hfill\begin{minipage}{0.45\textwidth}
\begin{lstlisting}[caption={OpenGL Layout},label={llst:opengl},language={GLSL},style={GL}]
//if SoA
#define Particle PREC_VEC_TYPE
#define Particle_position 0
#define Particle_velocity 1
#define Particle_size 2
//if AoS
struct Particle{
  PREC_VEC_TYPE position;
  PREC_VEC_TYPE velocity;
};
buffer Container{
  Particle variables [ ];
};
\end{lstlisting}
\end{minipage}
\end{wrapfigure}

@@

\subsection{Parallel Reduction}
\cite{NVIDIA:PARALLEL_REDUCTION}
\subsection{Counting sort}
Map (Reset) \\
\subsubsection{Bin}
Binning\cite{NVIDIA:BINNING} (coalesced, uncoalesced)\\
\subsubsection{Scan}
Scan\cite{NVIDIA:SCAN}\\
Scan(Raking) \cite{NVIDIA:SCAN_MODERN}
\subsubsection{Reordering}
Coalesced, Uncoalesced
\subsection{MPM-Transfers}
Conjugate gradient method
\subsubsection{Particle to grid transfers}
\subsubsection{Grid to particle transfers}
\subsection{Verifying Simulation}
\begin{itemize}
  \item Conservation equations
    \cite{MPM:APIC}
  \item Singular Value Decomposition
  \item Conjugate gradient
\end{itemize}
\clearpage
\section{Evaluation}
\textbf{SoA vs. AoS}:

\begin{table}
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Description &  $\Delta t_c (\mu s)$ & Speedup &VRAM     & SM     & L2     & Tex \\\hline
    AoS         & 278                   & -       &77.7\%   & 7.3\%  & 30.3\% & 14.7\% \\\hline
    SoA         & 122                   & 2.26x   &75.4\%   & 14.3\% & 29.4\% & 17.9\% \\
    \hline
  \end{tabular}
  \caption{Map operation with one instruction of form \texttt{float x = length2( (vec4) y)} on $1024 \times 1024$ \texttt{vec4}.}\label{table:map_1}
\end{table}

\begin{table}
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Description &  $\Delta t_c (\mu s)$ & Speedup &VRAM   & SM     & L2     & Tex \\\hline
    AoS         & 292                   & -       &61.3\% & 11.4\% & 41.8\% & 20.4\% \\\hline
    SoA         & 252                   & 1.16x   &75.4\% & 14.3\% & 29.4\% & 17.9\% \\
    \hline
  \end{tabular}
  \begin{tabular}{ | l | l | l | l | }    \hline
    Description &  L2-Hit & Read Active \\\hline
    AoS   & 53.8\%        & 48.9\%      \\\hline
    SoA   & 20.0\%          & 62.3\%      \\
    \hline
  \end{tabular}
  \caption{Map operation with two instructions of form \texttt{float x = length2( (vec4) y)} on $1024 \times 1024 \times 2$ \texttt{vec4}.}\label{table:map_2}
\end{table}

\begin{figure}[htbp]
  \centering
  \includesvg[svgpath =images/,width=\textwidth]{soa}
  \caption{SoA-Layout in practive. The first instruction is shown in orange, the second instruction in green. If green is left out throughput is still maximized}
\end{figure}
\begin{figure}[htbp]
  \centering
  \includesvg[svgpath =images/,width=\textwidth]{aos}
  \caption{AoS-Layout in practice. The first instruction is shown in orange, the second instruction in green. Orange already loads both cache-lines. If green is left out throughput is halved. Also green is an L2-cache lookup requiring additional overhead over the direct SoA way.}
\end{figure}
\section{Conclusion}
\subsection{Review}
\subsection{Future Work}
Particle Activation: \cite{MPM:GPU}
